---
title: "ETH24XX: Field Data Collection Tracker"
author: "REACH-Initiative"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(dirname(inputFile), '/eth_rna_field_dc_tracker_', format(Sys.Date(), '%Y_%m_%d'),'.html')) })
---

```{r logo, echo=FALSE}
htmltools::a(href = "https://www.reach-initiative.org",
             target = "_blank",
             htmltools::img(src = knitr::image_uri("../support_files/Logo_Reach_RGB_1.png"),
                            alt = "REACH logo",
                            style = 'position:absolute; top:-1; right:0; padding:0; margin:20; width:250px')
             )
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# read packages
library(tidyverse)
library(lubridate)
library(glue)
library(leaflet)

df_logical_check_description <- readr::read_csv("../support_files/rna_logical_check_list.csv") %>% 
    rename(check_number = check_id, check_description = description) %>% 
    janitor::clean_names() %>% 
    select(check_number, check_description) %>% 
    mutate(check_number = as.character(check_number))

df_samples <- readxl::read_excel("../support_files/rna_sample_size.xlsx")

df_for_colnames <- df_samples %>% 
    select(admin4) %>% 
    unique()

df_samples_required <- df_samples %>% 
    select(admin4, tot_sample_size) 

df_tool_data <- readxl::read_excel("../inputs/ETH24XX_RNA_Tigray_HH_data.xlsx") %>%
    mutate(uuid = `_uuid`,
           start_date = as_date(start),
           start = as_datetime(start),
           end = as_datetime(end),
           admin4 = as.character(admin4),
           latitude = as.numeric(`_hh_location_latitude`),
           longitude = as.numeric(`_hh_location_longitude`))

# days that contain data for refugees
df_days_for_data_collection <- df_tool_data %>% select(start_date) %>% unique() %>% arrange(start_date) %>% pull()

df_data_support_cl_log <- df_tool_data %>% 
  select(uuid, admin4, latitude,	longitude)

# cleaning log handling
df_cl_log <- readxl::read_excel("../inputs/main_combined_checks_eth_rna.xlsx") %>% 
    filter(!str_detect(string = admin4, pattern = fixed('test', ignore_case = TRUE))) %>% 
    mutate(change_type = ifelse(is.na(change_type), "apply_suggested_change", change_type)) %>% 
    left_join(df_data_support_cl_log, by = "uuid")

# change_response logs that affect stats in the data collection progress
cl_log_change_response <- df_cl_log %>% 
    filter(change_type == "change_response", 
           !is.na(new_value), reviewed == 1, 
           change_type != "no_action",
           !str_detect(string = question, "\\/")) %>% 
    select(uuid, question, new_value)

# updated tool data
df_updated_tool_data <- df_tool_data

# get uuids from cleaning log
uuids_chg_response <- cl_log_change_response %>% pull(uuid) %>% unique()

for (current_uuid in uuids_chg_response) {
    current_uuid_data <- cl_log_change_response %>% 
        filter(uuid == current_uuid) %>% 
        mutate(new_value = ifelse(question == "enumerator_id", as.numeric(new_value), new_value)) %>% 
        pivot_wider(names_from = "question", values_from = "new_value", uuid)
    print(current_uuid_data)
    # process current updates
    df_current_updated <- df_updated_tool_data 
    
    # update the parent dataset with current updates
    df_updated_tool_data <- df_current_updated
}

# enumerator performance data
df_enum_performance <- df_updated_tool_data %>% 
    mutate(int.survey_time_interval = lubridate::time_length(end - start, unit = "min"),
           int.survey_time_interval = ceiling(int.survey_time_interval))

# surveys for deletion
df_cl_deletion <- df_cl_log |> 
  filter(change_type %in% "remove_survey", reviewed == 1, change_type != "no_action") %>% 
  distinct(uuid)

# functions for changing some options in the table
dt_set_options<- function(x){
  DT::datatable(x,
                options = list(autoWidth = F,
                               dom= 't',
                               list(list(width = '20%', targets = list(1,2,3,4,5)))
                               )
                )
}

dt_with_modified_options <- function(x){
  DT::datatable(x,
                rownames = FALSE,
                options = list(
                  columnDefs = list(list(className = 'dt-center', targets = list(1,2,3,4,5))),
                  pageLength = 10,
                  initComplete = JS(
                    "function(settings, json) {",
                    "$(this.api().table().header()).css({'background-color': '#333', 'color': '#fff'});",
                    "}")
                )
  )
}

dt_options_fewcols <- function(x){
  DT::datatable(x,
                rownames = FALSE,
                options = list(
                  pageLength = 10,
                  initComplete = JS(
                    "function(settings, json) {",
                    "$(this.api().table().header()).css({'background-color': '#333', 'color': '#fff'});",
                    "}")
                )
  )
}

dt_enum_performance_options <- function(x){
  DT::datatable(x,
                rownames = FALSE,
                filter = 'top',
                options = list(
                  columnDefs = list(list(className = 'dt-center', targets = list(1,2))),
                  pageLength = 10,
                  initComplete = JS(
                    "function(settings, json) {",
                    "$(this.api().table().header()).css({'background-color': '#333', 'color': '#fff'});",
                    "}"),
                  order = list(list(1, 'desc'), list(0, 'asc'), list(3, 'desc'))
                )
  )
}
```

## Summary on the surveys done

>Total surveys: **`r df_updated_tool_data %>% nrow()`**,\
Surveys for deletion: **`r nrow(df_cl_deletion)`**,\
Last date of data collection: **`r df_days_for_data_collection[length(df_days_for_data_collection)]`**.

### Summary on the surveys per location

```{r, echo = FALSE}
df_sample_per_admin4 <- df_samples_required %>% 
    group_by(admin4) %>% 
    summarise(samples = sum(tot_sample_size, na.rm = TRUE))

df_cl_surveys_for_deletion <- df_cl_log %>% 
  filter(change_type == "remove_survey", change_type != "no_action", 
         reviewed == 1) %>% 
  rename(admin4 = admin4.x) %>% 
  group_by(admin4) %>% 
  distinct(uuid) %>%
  summarise(surveys_for_deletion = n())

  df_xt <- df_updated_tool_data %>% 
    group_by(admin4) %>% 
    summarise(surveys_completed = n()) %>% 
    arrange(admin4) %>% 
    right_join(df_sample_per_admin4, by = "admin4") %>% 
    left_join(df_cl_surveys_for_deletion, by = "admin4") %>% 
    mutate(surveys_completed = ifelse(is.na(surveys_completed), 0, surveys_completed),
           surveys_for_deletion = ifelse(is.na(surveys_for_deletion), 0, surveys_for_deletion),
           int.surveys_and_deletion = surveys_completed - surveys_for_deletion,
           remaining_surveys = samples - int.surveys_and_deletion) %>% 
    left_join(df_for_colnames, by = "admin4") %>% 
    select(-c(int.surveys_and_deletion)) 
  
  df_xt %>% 
    DT::datatable()
```

## Gender Disaggregation

<center>**Gender of Interview Tracking or Disaggregation** </center> </br> 

```{r, echo = FALSE}
# Gender disaggregation by respondent status
df_gender_disaggregation <- df_tool_data %>% 
    group_by(admin4) %>% 
    # Count occurrences of each gender within each admin4 region
    count(resp_gender = ifelse(is.na(hoh_gender), resp_gender, hoh_gender), name = "count") %>% 
    # Reshape the data to have separate columns for each gender
    pivot_wider(names_from = resp_gender, 
                values_from = count, 
                values_fill = 0) %>%  # Replace missing values with 0
    # Calculate gender percentages within each admin4 group
    mutate(male_percentage = male / (female + male) * 100,
           female_percentage = female / (female + male) * 100) %>% 
    # Round percentages to the nearest whole number
    mutate(across(ends_with("_percentage"), round, 0)) %>% 
    # Rename percentage columns for better readability
    rename(`male(%)` = male_percentage,
           `female(%)` = female_percentage)

# Display the results as an interactive table
df_gender_disaggregation %>% 
    DT::datatable()
```

### Daily enumerator performance

>The average survey time for all the data is: **`r round(mean(df_enum_performance$int.survey_time_interval), 0)`** Minutes

<center> **Number of survey per day of data collection, and average interview duration by enumerator** </center> </br>

```{r, echo = FALSE}
df_enum_performance %>% 
  group_by(admin4, start_date, enumerator_id) %>% 
  summarise(number_of_interviews_done = n(), `average_survey_time(minutes)` = round(mean(int.survey_time_interval, na.rm = TRUE), 0)) %>% 
  dt_enum_performance_options()
```

## Looking into the cleaning log:

### a) Number of issues by issue_id

<center> **Number of Flags by Enumerator** </center> </br>

```{r, echo = FALSE}
df_cl_log |> 
  group_by(issue) |> 
  summarise(number_of_issues_by_issue_id = n()) |>
  # mutate(int.issue_id = str_extract(string = issue_id, pattern = "[0-9]{1,3}")) |> 
  # left_join(df_logical_check_description, by = c("int.issue_id" = "check_number")) |> 
  # mutate(issue = ifelse(str_detect(string = issue_id, pattern = "[0-9]{1,3}"), check_description, issue_id)) |> 
  # select(-c(int.issue_id, check_description))  |> 
  dt_options_fewcols()
```

### b) Number of issues by enumerator

```{r, echo = FALSE}
df_cl_log %>% 
    rename(admin4 = admin4.x) %>% 
  group_by(admin4, enumerator_id) %>% 
  summarise(number_of_issues_by_enumerator_id = n()) %>%
  dt_options_fewcols()
```

### c) Number of issues by enumerator and issue_id

```{r, echo = FALSE}
df_cl_log |> 
  group_by(enumerator_id, issue) |> 
  summarise(number_of_issues_by_enumerator_and_issue_id = n()) |>
  # mutate(int.issue_id = str_extract(string = issue_id, pattern = "[0-9]{1,3}")) |> 
  # left_join(df_logical_check_description, by = c("int.issue_id" = "check_number")) |> 
  # mutate(issue = ifelse(str_detect(string = issue_id, pattern = "[0-9]{1,3}"), check_description, issue_id)) |> 
  # select(-c(int.issue_id, check_description))  |> 
  dt_options_fewcols()
```

### d) Enumerators with surveys for deletion

```{r, echo = FALSE}
df_cl_log %>% 
  filter(change_type == "remove_survey", 
         reviewed == 1) %>% 
    rename(admin4 = admin4.x) %>% 
  group_by(admin4, enumerator_id) %>% 
  summarise(number_of_surveys_for_deletion_by_enumerator = n()) %>%
  dt_options_fewcols()
```

### e) Map of Survey Distribution

```{r, echo = FALSE, out.width = "100%"}
# Create a leaflet map for geographic distribution
 data <- df_tool_data %>%
   mutate(latitude = as.numeric(`_hh_location_latitude`),
          longitude = as.numeric(`_hh_location_longitude`))

leaflet(data) %>%
  addTiles() %>%
  addMarkers(lng = ~longitude, lat = ~latitude, 
             popup = ~as.character(`_uuid`),
             clusterOptions = markerClusterOptions())
```

### f) Map of surveys for deletion

```{r, echo = FALSE, out.width="100%"}
# popup
 labels_pts <- ~sprintf(
   "<strong>Status and Name: %s</strong><br/>
       Issue :  <strong>%s</strong><br/>
       Enumerator ID :  <strong>%s</strong>",
   int.status, issue, enumerator_id) %>%
   lapply(htmltools::HTML)
 
 df_cl_log %>%
   filter(change_type == "remove_survey", reviewed == 1) %>%
     rename(admin4 = admin4.x) %>%
   group_by(uuid, admin4, latitude, longitude) %>%
   summarise(enumerator_id = paste(enumerator_id, collapse = " : "),
             admin4 = paste(admin4, collapse = " : "),
             #point_number = paste(point_number, collapse = " : "),
             change_type = paste(change_type, collapse = " : "),
             question = paste(question, collapse = " : "),
             new_value = paste(new_value, collapse = " : "),
             #issue_id = paste(issue_id, collapse = " : "),
             issue = paste(issue, collapse = " : ")) %>%
   unique() %>%
   mutate(int.status = glue("{admin4}_{admin4}")) %>%
   leaflet() %>%
   addTiles() %>%
   addCircleMarkers(~longitude,
                    ~latitude,
                    popup = labels_pts,
                    radius = 10,
                    color = "red",
                    stroke = FALSE, fillOpacity = 0.9,
                    label = labels_pts,
                    clusterOptions = markerClusterOptions())
```

## <strong> Feedback </strong>

This Data Collection Monitoring Report was generated on *`r format(Sys.Date(), format = "%B %d, %Y")`*. For any suggestions or improvements to this product, please feel free to reach out to: getu.gari@reach-initiative.org.